---
title: "Homework 1"
output: github_document
date: "2025-09-18"
author: Evan Kennedy, ELK2149
---

```{r homework_1_setup, include = FALSE}

#setting up
#opening necessary libraries
library(moderndive)
library(tidyverse)

#extracting data
data("early_january_weather")

#reviewing help file
help("early_january_weather")
```

**Problem 1**  

The variables in this dataset are: `r (names(early_january_weather))`.

Flights were recorded from early Jan `r unique(early_january_weather $ year)`. The average temperature across this span of time was `r mean(early_january_weather$temp)` °F.

The dataset has `r nrow(early_january_weather)` rows and `r ncol(early_january_weather)` columns.

The following is a scatter plot of temperature by date and hour across early Jan.

```{r tempXtime_hour_scatter, echo = FALSE}

#creates scatter plot of temp by time_hour
ggplot(early_january_weather, aes(x = time_hour, y = temp, color = humid)) + 
    geom_point()

#saves scatter plot
ggsave("tempXtime_hour_scatter.pdf", height = 4, width = 6)

```

Temperature and humidity appear to be increasing over time, with daily fluctuation, and humidity peaking around Jan 14th.

**Problem 2**  

The below code samples 10 variables from a normal distribution, evaluates how many of these variables are above zero, creates a character vector of length 10, and a factor vector of three different "levels".

```{r problem_2}

#Generating normal distribution, evaluating which are above 0, creating a character vector, and factor vector.

pb_2 = tibble(
  norm_samp = rnorm(10),
  norm_samp_pos = norm_samp > 0,
  vec_char = c("One", "Two", "Three", "Four", "Five", "Six", "Seven", "Eight", "Nine", "Ten"),
  vec_factor = factor(c("Level 1", "Level 1", "Level 1", "Level 2", "Level 2", "Level 2", "Level 3", "Level 3", "Level 3", "Level 3"))
)    

pb_2
```

The below code takes the mean of each variable.

```{r problem_2a}

mean_norm_samp <- pb_2 |> # mean of the variables
  pull(norm_samp) |>    # extracts the "norm_samp" column
  mean()

mean_norm_samp

mean_norm_samp_pos <- pb_2 |> # mean of how many are above 0
  pull(norm_samp_pos) |>    # extracts the "norm_samp_pos" column
  mean()

mean_norm_samp_pos

mean_vec_char <- pb_2 |> # attempted mean of the character vector
  pull(vec_char) |>    # extracts the "vec_char" column
  mean()

mean_vec_factor<- pb_2 |> # attempted mean fo the factor vector
  pull(vec_factor) |>    # extracts the "vec_factor" column
  mean()


```

You cannot take a mean of the non-numeric character and factor variables, so it returns as NA. The mean of mean_norm_samp_pos is actually the proportion of variables that are above zero.

```{r problem_2b, eval = FALSE}

mean_norm_samp_pos <- pb_2 |> 
  pull(norm_samp_pos) |>    # extracts the "norm_samp_pos" column
  as.numeric() |>   # converts to numeric value
  mean()

mean_vec_char <- pb_2 |> 
  pull(vec_char) |>    # extracts the "vec_char" column
  as.numeric() |>   # converts to numeric value
  mean()

mean_vec_factor<- pb_2 |> 
  pull(vec_factor) |>    # extracts the "vec_factor" column
  as.numeric() |>     # converts to numeric value
  mean()

```

You can use as.numeric to convert factors, and some characters into numeric values. Because I spelled out my character numbers (one - ten), rather than using 1-10, this didn't work for the character vector.
